---
title: "NextGen forcing analysis CSV to NetCDF conversion"
author: "Keith Jennings"
date: "9/27/2022"
output: github_document
---

A quick look at NetCDF vs CSV forcing for NextGen

Punch line: they're both the same, except for slight far decimal differences.

The below code imports, matches, then analyzes the data.

```{r}

# Script for converting NLDAS-2 NETCDF files to CSV
# Data downloaded from NASA GES DISC
# Data are for Irwin Creek, catchment 87

# Keith Jennings
# keith.jennings@noaa.gov
# 2020-11-20

#Load packages
library(tidyverse)
library(ncdf4)

# Identify input and output directories
input_dir = "../data/forcing/"

# list files
nc_files <- data.frame(files = list.files(path = input_dir, pattern = ".nc$"), 
                       stringsAsFactors = F)
csv_files <- data.frame(files = list.files(path = input_dir, pattern = ".csv$"),
                        stringsAsFactors = F)

# Open the NetCDF file
nc <- nc_open(paste0(input_dir, nc_files[1, "files"]))

# Read and concatenate all of the csv data
df = data.frame()
ncRAINRATE = ncvar_get(nc, "RAINRATE")
ncT2D = ncvar_get(nc, "T2D")
ncQ2D = ncvar_get(nc, "Q2D")
ncU2D = ncvar_get(nc, "U2D")
ncV2D = ncvar_get(nc, "V2D")
ncPSFC = ncvar_get(nc, "PSFC")
ncSWDOWN = ncvar_get(nc, "SWDOWN")
ncLWDOWN = ncvar_get(nc, "LWDOWN")
ncIDS = ncvar_get(nc, "ids")

for(i in 1:length(csv_files$files)){
  # identify the catchment
  tmpCat = str_extract(csv_files[i, "files"], "[^.]+")
  
  # Get the numeric column corresponding to the catchment in the netcdf file
  tmpIds = which(ncIDS == tmpCat)
 
  # Import the csv and add netcdf data for that catchment
  tmp <- read.csv(paste0(input_dir, csv_files[i, "files"])) %>% 
     mutate(cat = tmpCat,
         ncRAINRATE = ncRAINRATE[, tmpIds],
         ncT2D      = ncT2D[, tmpIds],
         ncQ2D      = ncQ2D[, tmpIds],
         ncU2D      = ncU2D[, tmpIds],
         ncV2D      = ncV2D[, tmpIds],
         ncPSFC     = ncPSFC[, tmpIds],
         ncSWDOWN   = ncSWDOWN[, tmpIds],
         ncLWDOWN   = ncLWDOWN[, tmpIds],)
 
  # Bind to dataframe
  df <- bind_rows(df, tmp)
}

# Make data longer
df_l <- 
  bind_cols(
    df %>% select(RAINRATE:cat) %>% 
      pivot_longer(where(is.numeric), names_to = "var_csv", values_to = "value_csv"),
    df %>% select(ncRAINRATE:ncLWDOWN) %>% 
      pivot_longer(where(is.numeric), names_to = "var_nc", values_to = "value_nc")
  )

# Summarize the data for bias
df_summary <- df_l %>% 
  na.omit() %>% 
  group_by(cat, var_csv) %>% 
  summarise(mean_bias = mean(value_csv - value_nc),
            corr = cor(value_csv, value_nc))

```

Here's a table of per-catchment and variable biases and correlations:

```{r echo = FALSE, warning = FALSE, message = FALSE}
df_summary %>% 
  knitr::kable()
```

Most biases are 0 or near 0 and all correlations are 1. For non-zero biases, the difference appears to be at a very distant decimal point:

```{r}
df %>% slice(1) %>% pull(ncT2D) %>% sprintf("%.10f", .)
df %>% slice(1) %>% pull(T2D) %>% sprintf("%.10f", .)
```

Probably an error induced by putting the CSV data into the NetCDF file.